A Theory of Bidirectional Type Synthesis for Simple Types

### 2023-06-29 tex/sec2-outline.tex
[ tex ] Fine-tune the figure of overview

\begin{figure}
  \small
  \bgroup
  \renewcommand{\arraystretch}{1.5}
  \begin{tabular}{ r l }
    $\boxed{\Gamma |- \isTerm{t} \syn A}$ & A raw term $t$ synthesises a type $A$ under $\Gamma$ \\
    $\boxed{\Gamma |- \isTerm{t} \chk A}$ &A raw term $t$ checks against a type $A$ under $\Gamma$
  \end{tabular}
  \egroup
  \centering
  \begin{mathpar}
    \inferrule{(\isTerm{x} : A) \in \Gamma}{\Gamma \vdash \isTerm{x} \syn A}\,\SynRule{Var}
    \and
    \inferrule{\Gamma \vdash \isTerm{t} \chk A}{\Gamma \vdash \isTerm{(t \annotate A)}\syn A}\,\SynRule{Anno}
    \and
    \inferrule{\Gamma \vdash \isTerm{t} \syn A \\ A = B}{\Gamma \vdash \isTerm{t} \chk B}\,\ChkRule{Sub}
    \and
    \inferrule{\Gamma, \isTerm{x} : A \vdash \isTerm{t} \chk B}{\Gamma \vdash \isTerm{\lam{x}t} \chk A \bto B}\,\ChkRule{Abs}
    \and
    \inferrule{\Gamma \vdash \isTerm{t} \syn A \bto B \\ \Gamma \vdash \isTerm{u} \chk A}{\Gamma \vdash \isTerm{t\;u} \syn B}\,\SynRule{App}
  \end{mathpar}
  \caption{Bidirectional typing derivations for simply typed $\lambda$-calculus}
  \label{fig:STLC-bidirectional-typing-derivations}
\end{figure}

@@

\begin{figure}
  \begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=3em, column sep=10em]
  {
    \boxed{\erase{\Gamma} \vdash \isTerm{t}}   & \boxed{\erase{\Gamma} \vdash \isTerm{t}^{\syn}} \\
    \boxed{\Gamma \vdash \isTerm{t} : A} & \boxed{\Gamma \vdash \isTerm{t} \syn A} \\
  };
  \path[->]
      (m-1-1) edge node [above,align=center] {mode preprocessing} (m-1-2)
              edge node [left,align=left]  {type\\synthesis} (m-2-1)
      (m-1-2) edge node [right,align=right] {bidirectional\\type synthesis} (m-2-2)
      (m-2-1) edge[bend right=8] node [below] {completeness} (m-2-2);
  \path[->]
      (m-2-2) edge[bend right=8] node [above] {soundness} (m-2-1);
  \end{tikzpicture}
  \caption{An implementation of type synthesis via bidirectional typing}
\end{figure}

>>>>

\begin{figure}
  \small
  \bgroup
  \renewcommand{\arraystretch}{1.5}
  \begin{tabular}{ r l }
    $\boxed{\Gamma |- \isTerm{t} \syn A}$ & A raw term $t$ synthesises a type $A$ under $\Gamma$ \\
    $\boxed{\Gamma |- \isTerm{t} \chk A}$ &A raw term $t$ checks against a type $A$ under $\Gamma$
  \end{tabular}
  \egroup
  \centering
  \begin{mathpar}
    \inferrule{(\isTerm{x} : A) \in \Gamma}{\Gamma \vdash \isTerm{x} \syn A}\,\SynRule{Var}
    \and
    \inferrule{\Gamma \vdash \isTerm{t} \chk A}{\Gamma \vdash \isTerm{(t \annotate A)}\syn A}\,\SynRule{Anno}
    \and
    \inferrule{\Gamma \vdash \isTerm{t} \syn B \\ B = A}{\Gamma \vdash \isTerm{t} \chk A}\,\ChkRule{Sub}
    \and
    \inferrule{\Gamma, \isTerm{x} : A \vdash \isTerm{t} \chk B}{\Gamma \vdash \isTerm{\lam{x}t} \chk A \bto B}\,\ChkRule{Abs}
    \and
    \inferrule{\Gamma \vdash \isTerm{t} \syn A \bto B \\ \Gamma \vdash \isTerm{u} \chk A}{\Gamma \vdash \isTerm{t\;u} \syn B}\,\SynRule{App}
  \end{mathpar}
  \caption{Bidirectional typing derivations for simply typed $\lambda$-calculus}
  \label{fig:STLC-bidirectional-typing-derivations}
\end{figure}

@@

\begin{figure}
  \begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=3em, column sep=10em]
  {
    \boxed{\erase{\Gamma} \vdash \isTerm{t}}   & \boxed{\erase{\Gamma} \vdash \isTerm{t}^{\syn}} \\
    \boxed{\Gamma \vdash \isTerm{t} : A} & \boxed{\Gamma \vdash \isTerm{t} \syn A} \\
  };
  \path[->, line width=0.7pt]
      (m-1-1) edge node [above] {mode preprocessing} (m-1-2)
      (m-1-2) edge node [right,align=center] {bidirectional\\type synthesis} (m-2-2)
      (m-2-1) edge[bend right=8] node [below] {completeness} (m-2-2)
      (m-2-2) edge[bend right=8] node [above] {soundness}    (m-2-1);
  \path[dashed,->,line width=0.7pt]
      (m-1-1) edge node [left,align=center]  {type\\synthesis} (m-2-1);
  \end{tikzpicture}
  \caption{An implementation of type synthesis via bidirectional typing}
\end{figure}
